Combined effects of protein expression variance and correlation on multicomponent systems

### 2020-03-10 Manuscript/Manuscript.Rmd
.

```{r figure_2,, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_2}
allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 10)) %>%
  mutate(interval = ifelse(interval == levels(interval)[1],
                           "low",
                           ifelse(interval == levels(interval)[10],
                                  "high",
                                  "mid")))

ggplot(allGroups, aes(x = normVar)) + geom_histogram(bins = 100, aes(fill = interval)) + theme_minimal() +
  #scale_fill_viridis_d()
  scale_fill_brewer(palette = "PiYG")

ggplot(allGroups, aes(x = medCors, y = medCV, color = interval)) + 
  geom_point() + 
  theme_minimal()

#fig_legend_2 <- c("A) ")
```

>>>>

```{r figure_2,, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_2}
allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
  mutate(interval = ifelse(interval == levels(interval)[1],
                           "low",
                           ifelse(interval == levels(interval)[7],
                                  "high",
                                  "mid")))

p1 <- ggplot(allGroups, aes(x = normVar, fill = interval)) + 
  geom_histogram(bins = 50) + theme_minimal() +
  #scale_fill_viridis_d()
  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
  xlab("Normalized Variation") + ylab("Number")

interesting_groups_low <- c("03010","00030","GO:0046496","00071","GO:0006096","GO:0015934","GO:0019843","00051")
interesting_groups_high <- c("GO:0042995","03008","GO:0007264","GO:0005885","GO:0007166","04150","GO:0035556","GO:0016791","GO:0007165")
interesting_groups <- c(interesting_groups_low, interesting_groups_high)

p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval)) + 
  geom_point(data = filter(allGroups, interval == "mid", n > 5),
             size = 2, alpha = 0.5, stroke = 0) +
  geom_point(data = filter(allGroups, interval != "mid", n > 5),
             size = 3, alpha = 0.75, stroke = 0) +
  geom_label_repel(data = filter(allGroups, accession %in% interesting_groups), 
                  aes(label = name), 
                  colour = 'black', 
                  size = 2,
                  force = 100,
                  box.padding   = 1.5) +
  theme_minimal() +
  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
  xlab("Mean Correlation") + ylab("Mean Coefficient of Variation")

p3 <- ggplot(allGroups, aes(x = medCors, fill = interval)) + 
  annotate("rect", xmin = 0.15, xmax = Inf, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.5) + 
  geom_density(data = filter(allGroups, interval != "mid"), alpha = 0.75, color = NA) +
  geom_density(data = filter(allGroups, interval == "mid"), alpha = 0.5, color = NA) +
  theme_minimal() +
  scale_fill_manual(values = c("#d53e4f","#3288bd","grey"))

p4 <- ggplot(allGroups, aes(x = medCV, fill = interval)) + 
  geom_density(alpha = 0.75, color = NA) +
  theme_minimal() +
  scale_fill_manual(values = c("#d53e4f","#3288bd","grey"))

(p1 + p3 + p4) / p2 + plot_layout(heights = c(1,5)) +
  plot_annotation(tag_levels = "A") & 
  theme(legend.position = "none")

fig_legend_2 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways. The bottom and top 10% are colored blue and red respectively. B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A. Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance. In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
```

```{r figure_3, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_3}
glyc <- filter(allGroups, accession == '00010')$Uniprot %>% unlist()

glycMat <- cor[rownames(cor) %in% glyc,colnames(cor) %in% glyc]
rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')

row_anno <- data.frame(cv = unlist(filter(allGroups, accession == "00010") %>% pull(cv)))
rownames(row_anno) <- rownames(glycMat)

map <- pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
         cutree_cols = 3, cutree_rows = 3, annotation_row = row_anno)

map.clust <- cbind(t(glycMat), cluster = cutree(map$tree_row, k = 3))
map.clust[,ncol(map.clust) == 2]

glyc2 <- names(which(map.clust[,ncol(map.clust)] == 2))
totVarGroup(glyc2) / length(glyc2)
protVar %>% filter(Uniprot %in% glyc2) %>%
  summarise(meanCV = mean(cv))

glyc2 <- glycMat[which(map.clust[,ncol(map.clust)] == 2),which(map.clust[,ncol(map.clust)] == 2)]

# p7 <- as.ggplot(
#   Heatmap(cor, 
#         col = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201)),
#         heatmap_legend_param = list(title = "R"),
#         show_column_dend = FALSE,
#         show_row_dend = FALSE,
#         show_column_names = FALSE,
#         show_row_names = FALSE,
#         use_raster = TRUE,
#         raster_quality = 1)
#   )

p3 <- as.ggplot(
  pheatmap(glyc2, 
           color = mypal, 
           breaks = seq(-1,1, by = 0.01), 
           border_color = NA, 
           treeheight_col = 0,
           treeheight_row = 0,
           cellheight = 10, 
           cellwidth = 10,
           legend = FALSE)
)

glyc2 <- tibble(GO = "Glycolysis",
       cor = glyc2[lower.tri(glyc2)])

p1 <- rbind(glyc2,tibble(GO = "All",
      cor = cor[lower.tri(cor)])) %>%
  ggplot(.,aes(x = cor, fill = GO)) + geom_density(alpha = 0.75, color = "transparent") +
  scale_fill_manual(values = c('#bababa','#e08214')) +
  geom_rug(data = glyc2, color = '#e08214') +
  theme_bw() +
  #theme(legend.position = c(.2,.8)) +
  theme(legend.position = "none") +
  xlim(-1,1)

p2 <- rbind(glyc2,tibble(GO = "All",
      cor = cor[lower.tri(cor)])) %>%
  ggplot(., aes(x = GO, y = cor, fill = GO)) + 
  geom_boxplot(notch = TRUE, alpha = 0.75) +
  scale_fill_manual(values = c('#bababa','#e08214')) +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylim(-1,1) +
  coord_flip()

p4 <- p1/p2 + plot_layout(heights = c(2,1))

fig_legend_3 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways. The bottom and top 10% are colored blue and red respectively. B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A. Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance. In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
```
